### 2023-12-26 ms.qmd
Updates

Cognition is defined as the processes by which animals gather, preserve, and use information from their environment through perception, learning, memory, and decision making (@shettleworth). These mechanisms underpin several aspects of animals' ecology such as foraging, mate choice, antipredatory strategies, and/or social behaviours that are crucial for survival and reproduction (@dukas_evolutionary_2004). Particularly, learning - the acquisition of neuronal representations of new information (@dukas_evolutionary_2004) - is seen as fundamental for coping with environmental changes by allowing individuals to create new associations between events (@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013). However, the ability of organisms to acquire new information varies accross individuals, with this natural variation usually being influenced by factors like age, sex, gut microbiota, or the environment where animals develop (@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014). In this regard, the latter can be particularly relevant because the brain is especially sensitive to environmental impacts during early stages of development (@zhu_prenatal_2004). 
Prenatal Glucocorticoids (GCs) and prenatal thermal environment are known to play a prominent role in shaping learning abilities in different taxa (see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @crino_corticosterone_2014; @abayarathna_effects_2020). GCs - hormones related to organisms' response to stress (@sapolsky_how_2000) - can have sustained effects on neural structure and physiology that are associated with animals' performance on learning tasks (@lemaire_prenatal_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016). Some studies have demonstrated that prenatal stress and high prenatal GC levels impair with learning (@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn), while others revealed that these effects may vary depending on aspects like sex or learning task (@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016). Similarly, diverse experiments have shown pivotal impacts of prenatal temperature on learning in ectotherms (@amiel_hotter_2012; @amiel_egg_2014;). For instance, high incubation temperatures have been linked with faster learning rates in skinks (@amiel_hotter_2012; @amiel_egg_2014), but velvet geckos incubated at temperatures over their natural range learn slower than those incubated within the natural thermal limits (@abayarathna_effects_2020). In this vein, the effect of prenatal temperature has been linked to alterations in neural structure and metabolic activity (@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017; @beltran_are_2021) that are similar to those resulting from prenatal increased stress or GC levels (@lemaire_prenatal_2000; @zhu_prenatal_2004; @du_dynamic_2009). This suggests that prenatal GCs and temperature can act on the same physiological mechanisms and, as such, both may interact to shape individual variation on learning skills (@amiel_effects_2017; @noble_developmental_2018). Furthermore, GCs can play an important role in determining vertebrate responses to elevated temperatures (@Crino_2023) potentially driving to natural interactions between temperature and GCs. However, despite the similar mechanistic effects of prenatal GCs and temperature and the potential function of GCs in vertebrates response to elevated temperatures, our understanding of how these two factors interplay remains incomplete.
In this study, we aim to explore the interactive effects between prenatal Glucocorticoids (GCs) and the prenatal thermal environment on learning using two species of skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_), as our model species. We experimentally increased Corticosterone (CORT) - the main GC in birds, reptiles, amphibians, and rodents (@Crino_2023) - levels in the eggs of these two species of skinks and then incubated them at two different temperatures in a 2X2 factorial design. Then, we subjected the juveniles to a colour-associative and a reversal task to explore their learning abilities. We hypothesised that changes in CORT levels and temperature during early development would have sustained effects on brain's morphology and physiology that would ultimately impact learning skills. We predicted that individuals exposed to high levels of CORT and/or low temperatures would perform worse in the learning tasks than control individuals or individuals exposed to high temperatures. We also expected incubation at high temperatures to decrease the impact of CORT on skink's performance, while cold incubation temperatures are expected to increase CORT detrimental effects on learning.

>>>>

Cognition is defined as the processes by which animals gather, preserve, and use information from their environment through perception, learning, memory, and decision making (@shettleworth). These cognitive processes underpin several aspects of animals' ecology such as foraging, mate choice, antipredatory strategies, and/or social behaviours, all crucial for the survival and reproduction of animals (@dukas_evolutionary_2004). Notably, learning - the acquisition of neuronal representations of new information (@dukas_evolutionary_2004) - is seen as fundamental for coping with environmental changes by enabling individuals to create new associations between events (@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013). However, the capacity of individuals to acquire new information exhibits natural variation influenced by factors like age, sex, gut microbiota, or the environment where animals develop (@szuran_water_1994; @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @carazo_sex_2014; @noble_age-dependent_2014; @alemohammad_2022_microbiota_learning). The developmental environment, in particular, plays a pivotal role, as the brain is especially susceptible to environmental influences during early stages of development (@zhu_prenatal_2004). 
Prenatal Glucocorticoids (GCs) and prenatal thermal environment are known to play a prominent role in shaping learning abilities in different taxa (see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @crino_corticosterone_2014; @abayarathna_effects_2020). GCs - hormones related to organisms' response to stress (@sapolsky_how_2000) - exert sustained effects on neural structure and physiology that are associated with animals' performance on learning tasks (@lemaire_prenatal_2000; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016). Some studies have demonstrated that prenatal stress and high prenatal GC levels impair with learning (@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn), while others showed diverse effects depending on factors such as sex or the nature of the learning task (@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016). 
Similarly, some experiments have shown significant impacts of prenatal temperature on learning in ectotherms (@amiel_hotter_2012; @amiel_egg_2014;). For instance, high incubation temperatures have been linked with faster learning rates in skinks (@amiel_hotter_2012; @amiel_egg_2014), while velvet geckos incubated at temperatures over their natural range learn slower than those incubated within the natural thermal limits (@abayarathna_effects_2020). In this vein, the effect of prenatal temperature appears to be linked to alterations in neural structure and metabolic activity (@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017; @beltran_are_2021) that are similar to those resulting from prenatal increased stress or GC levels (@lemaire_prenatal_2000; @zhu_prenatal_2004; @du_dynamic_2009). This suggests that prenatal GCs and temperature can act on the same physiological mechanisms and, thus, both could interact to shape individual variation in learning skills (@amiel_effects_2017; @noble_developmental_2018). Furthermore, GCs can play an pivotal role in determining vertebrate responses to elevated temperatures (@Crino_2023) potentially fostering natural interactions between temperature and GCs. Despite the proximate similarities of prenatal GCs and temperature effects and the potential role of GCs in vertebrates response to elevated temperatures, our understanding of how these two factors interact remains incomplete.
In this study, our objective is to explore the interactive effects between prenatal Glucocorticoids (GCs) and the prenatal thermal environment on learning. We utilized two species of skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_), as model species. We experimentally increased Corticosterone (CORT) - the main GC in birds, reptiles, amphibians, and rodents (@Crino_2023) - levels in the eggs of these two species of skinks and then incubated them at two different temperatures in a 2X2 factorial design. Post-incubation, the juveniles were subjectd to a colour-associative and a reversal task to comprehensively assess their learning abilities. Our hypothesis posits that changes in CORT levels and temperature during early development will induce sustained effects on brain's morphology and physiology that will ultimately impact learning skills. We predict that individuals exposed to high levels of CORT and/or low temperatures will perform less proficiently in the learning tasks compared to control individuals or those exposed to high temperatures. Additionally, we anticipate that incubation at high temperatures will mitigate the impact of CORT on skink performance, while cold incubation temperatures are expected to enhance the detrimental effects of CORT on learning.
